gij bent 2

### 2025-10-01 paper/paper.qmd
Finish Julie figure

```{r adopter-categories}
#| echo: FALSE

# TODO volgorde legende is fucked

# Load libraries
library(ggplot2)

# Simulated time steps
time <- seq(0, 20, 0.1)

@@

# Combine into dataframe
df <- data.frame(time, s_curve, bell_curve)

# Define cutoffs for adopter categories
mean_time <- 10
sd_time <- 2.5
cuts <- data.frame(
  x = c(mean_time - 2*sd_time, mean_time - sd_time, mean_time,
        mean_time + sd_time, mean_time + 2*sd_time),
  label = c("Innovators", "Early adopters", "Early majority", 
            "Late majority", "Laggards"),
  col = c("red", "orange", "green3", "darkgreen", "blue")
)

# Plot
ggplot(df, aes(x = time)) +
  geom_line(aes(y = s_curve, colour = "S-curve (cumulative adoption)"), linewidth = 1) +
  geom_line(aes(y = bell_curve, colour = "Bell Curve (adoption rate)"), linewidth = 1, linetype = "dashed") +
  geom_vline(data = cuts, aes(xintercept = x, colour = label), linetype = "dashed") +
  scale_colour_manual(
    name = "Legend",
    values = c("S-curve (cumulative adoption)" = "purple",
               "Bell curve (adoption Rate)" = "deeppink",
               "Innovators" = "red",
               "Early adopters" = "orange",
               "Early majority" = "goldenrod3",
               "Late majority" = "darkgreen",
               "Laggards" = "blue")
  ) +
  labs(
    title = "S-curve and bell curve",
    x = "Time step",
    y = "Adoption percentage"
  ) +
  theme_minimal(base_size = 14)

>>>>

```{r adopter-categories}
#| echo: FALSE
#| fig-cap: S-curve showing the cumulative adoption or diffusion of an innovation. The bell curve is its derivative and shows the adoption rate over time. For *gij bent*, we are assumed to currently be in the "Early adopters" phase.

# Rescale time to 0–100
time <- seq(0, 20, 0.1)
time_rescaled <- time * 5   # 0–100

@@

# Combine into dataframe with rescaled x
df <- data.frame(time = time_rescaled, s_curve, bell_curve)

# Build cutoff bands instead of single lines
bands <- data.frame(
  xmin = c(0, 
           2.5, 
           16, 
           50, 
           84),
  xmax = c(2.5, 
           16, 
           50, 
           84, 
           100),
  label = c("Innovators", "Early adopters", "Early majority", 
            "Late majority", "Laggards"),
  fill = c("red", "orange", "yellow", "darkgreen", "blue")
)

# Ensure correct order
bands$label <- factor(
  bands$label,
  levels = c("Innovators", "Early adopters", "Early majority", "Late majority", "Laggards")
)

# Plot with bands
ggplot(df, aes(x = time)) +
  # bands (slightly transparent so curves still visible)
  geom_rect(data = bands, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
            alpha = 0.1, inherit.aes = FALSE) +

  geom_line(aes(y = s_curve, colour = "S-curve (cumulative adoption)"), linewidth = 1) +
  geom_line(aes(y = bell_curve, colour = "Bell curve (adoption rate)"), linewidth = 1, linetype = "dashed") +
  # label
  annotate("text", x = 9.25, y = 52, label = "Now",
           colour = "#7f6c4c", hjust = 0.5) +
  scale_fill_manual(
    name = "Adopter categories",
    values = c("Innovators" = "red",
               "Early adopters" = "orange",
               "Early majority" = "yellow",
               "Late majority" = "darkgreen",
               "Laggards" = "blue")
  ) +
  scale_colour_manual(
  name = "",
  values = c("S-curve (cumulative adoption)" = "purple",
             "Bell curve (adoption rate)" = "deeppink"),
  limits = c("S-curve (cumulative adoption)",
             "Bell curve (adoption rate)")) +  # enforce order
  labs(
    x = "Time",
    y = "Adoption"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  guides(
    colour = guide_legend(nrow = 1),  # lines legend
    fill   = guide_legend(nrow = 2)   # bands legend
  )
```
